---
title: "skript"
output: html_document
---

```{r setup, include=FALSE, cache = TRUE}
data <- tidytuesdayR::tt_load("2021-04-27")
```

## Tidy Tuesday

This is my third contribution to TidyTuesday, which is 'a weekly podcast and community activity brought to you by the R4DS Online Learning Community'. Their goal is to help R learners learn in real-world contexts. 

For more information, visit the [TidyTuesday homepage](https://www.tidytuesday.com/), check out their [GitHub repository](https://github.com/rfordatascience/tidytuesday) and follow the [R4DS Learning Community on Twitter](https://twitter.com/R4DScommunity).

The purpose of these posts is mainly for exercising purposes. Thus, the provided graphs are not necessarily designed to provide the greatest possible insights. However, I always provide the R code for interested people at the page bottom.

```{r, include=FALSE}
library(tidyverse)
library(showtext)
library(ggtext)
library(stringr)
library(genderizeR)
library(gender)
knitr::opts_chunk$set(fig.width=12, fig.height=6) 
ceos <- data$departures
# use only first word to extract first names
ceos$fnames <- word(string = ceos$exec_fullname, start = 1, end = 1, sep = fixed(" "))
# for some, first word is abbrev.; take the second word as first name
ceos$fnames[nchar(ceos$fnames) == 2] <- word(string = ceos$exec_fullname[nchar(ceos$fnames) == 2], start = 2, end = 2, sep = fixed(" "))
# sort out CEOs with more than one abbrev first name
ceos <- ceos[nchar(ceos$fnames) > 2, ]
gender1 <- gender(ceos$fnames[1:1000], method = "genderize")
gender2 <- gender(ceos$fnames[1001:2000], method = "genderize")
gender3 <- gender(ceos$fnames[2001:3000], method = "genderize")
gender4 <- gender(ceos$fnames[3001:4000], method = "genderize")
gender5 <- gender(ceos$fnames[4001:5000], method = "genderize")
```

## References 

- `genderizeR`**-package** :  
Wais K (2006). "Gender Prediction Methods Based on First Names with genderizeR." _The R Journal_, *8*(1), 17-37. doi: [10.32614/RJ-2016-002](http://doi.org/10.32614/RJ-2016-002)

--------------------------------------------------

Full R code available at <https://github.com/jgranna/tidytuesday>