---
title: "Managers"
output: html_document
---


```{r setup, include=FALSE, cache = TRUE}
data <- tidytuesdayR::tt_load("2021-05-18")
```

## Tidy Tuesday

This is my fifth contribution to TidyTuesday, which is 'a weekly podcast and community activity brought to you by the R4DS Online Learning Community'. Their goal is to help R learners learn in real-world contexts. 

For more information, visit the [TidyTuesday homepage](https://www.tidytuesday.com/), check out their [GitHub repository](https://github.com/rfordatascience/tidytuesday) and follow the [R4DS Learning Community on Twitter](https://twitter.com/R4DScommunity).

The purpose of these posts is mainly for exercising purposes. Thus, the provided graphs are not necessarily designed to provide the greatest possible insights. However, I always provide the R code for interested people at the page bottom.

```{r, include=FALSE}
library(tidyverse)
library(showtext)
library(ggtext)
library(ggforce)
knitr::opts_chunk$set(fig.width=12, fig.height=6) 

managers <- data$survey
```

## Ask a Manager Survey

This week's data comes from [Ask a Manager Survey](https://docs.google.com/spreadsheets/d/1IPS5dBSGtwYVbjsfbaMCYIWnOuRmJcbequohNxCyGVw/edit?resourcekey#gid=1625408792). 

### Scatterplot of the data

A scatterplot of the data shows the locational distribution of the observations. Arrows indicate countries with a relatively high water source density.

```{r, include=FALSE}
# filter out 5% highest and lowest salaries
managers_ref <- managers
managers <- managers %>% 
  filter(annual_salary < quantile(managers_ref$annual_salary, probs = 0.95)) %>%
  filter(annual_salary > quantile(managers_ref$annual_salary, probs = 0.05))
experience <- c("1 year or less", "2 - 4 years", "5-7 years", "8 - 10 years", "11 - 20 years", "21 - 30 years", "31 - 40 years", "41 years or more")
managers$years_of_experience_in_field <- factor(managers$years_of_experience_in_field, levels = experience, labels = experience)
```
```{r, echo=FALSE, warning=FALSE, message=FALSE}
font_add_google("Open Sans", "osans")
showtext_auto()
theme_set(theme_minimal(base_size = 30, base_family = "osans"))
g1 <- ggplot(managers, aes(x = years_of_experience_in_field, y = annual_salary)) +
  geom_violin(fill = "white") +
  geom_sina(alpha = 0.1)  +
  labs(
    title = "**Manager Annual Salaries versus Experience**",
    subtitle = "Violin plot of bivariate relationship",
    caption = "**Data:** Ask a Manager Survey | **Visualization:** @jugRanna",
    x = "Years of experience",
    y = "annual salary"
  ) +
  theme(
    plot.title.position = "plot",
    plot.title = element_markdown(size = 43),
    axis.title.x = element_markdown(),
    axis.title.y = element_markdown(),
    plot.caption = element_markdown(color = "gray70")
  ) 
g1
```

## References

- `ggtext`**-package**:  
<https://wilkelab.org/ggtext/>

- `ggforce`**-package**:  
<https://ggforce.data-imaginist.com/>

- `showtext`**-package**:  
<https://github.com/yixuan/showtext>

- `tidyverse`**-package**:  
<https://www.tidyverse.org/>


--------------------------------------------------

Full R code available on [Github](https://github.com/jgranna/tidytuesday/tree/main/2021-05-04).