---
title: "Canadian Heatwave"
output: html_document
bibliography: bibliography.bib
---


```{r setup, include=FALSE, cache = TRUE}
library(tidyverse)
library(showtext)
library(ggtext)
library(colorspace)
## data
## via https://projects.oregonlive.com/weather/temps/
df <- 
  readr::read_csv("https://projects.oregonlive.com/weather/pdx_temps.csv") %>% 
  mutate(yday = lubridate::yday(date),
         year = lubridate::year(date),
         years = year %/% 20 * 20,
         decade = year %/% 10 * 10,
         decade = if_else(decade == 2020, "2020–2021", paste0(decade, "–", decade + 9)),
         years = if_else(years == 2020, "2020–2021", paste0(years, "–", years + 19)),
         new_record = if_else(tmax > 100 & year == 2021, "Hottest 3 Days ever", ""), #New heat records in 2021
         desc = "June 26 - June 28 were the hottest three days ever recorded in Portland.")

months <- tibble(yday = c(1, 32, 60, 91, 121, 152, 182, 213, 244, 274, 305, 335),
                 label = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", " ", "Sep", "Oct", "Nov", "Dec"))
```

Regarding the exceptional heatwave in Canada in June 2021, I decided to pause TidyTuesday for a week and instead visualize minimum and maximum temperatures in Portland between 1938 and 2021. This work is inspired by a post by the [New York Times](https://twitter.com/DrShepherd2013/status/1409911531052146690/photo/1) and a tweet by [Cedric Scherer](https://twitter.com/CedScherer/status/1410180270930731009/photo/1).

```{r, include=FALSE}
library(tidyverse)
library(showtext)
library(ggtext)
library(colorspace)
knitr::opts_chunk$set(fig.width=8, fig.height=5) 
```

## Scatter Plot of Maximum Temperatures in Portland, Canada

June 26-28, 2021, were the hottest days ever recorded in Portland, Canada, reaching a maximum of 116 degrees Fahrenheit (roughly 46.7 degrees Celsius). To show how unusually hot these days are in the last roughly 80 years, I provide a scatter plot, where the outlier temperatures are easily identifyable. I choose a plot on a polar axis and to indicate possible time trends, I color the points corresponding to their respective 20-year period. I also color the main y-grid lines for slightly improved interpretability.

```{r, echo = FALSE, warning=FALSE}
font_add_google("Roboto Condensed", "osans")
showtext_auto()
theme_set(theme_void(base_family = "osans"))
pal1 <- sequential_hcl(5, palette = "Red-Blue", rev = T)
theme_update(
  plot.caption = element_markdown(size = 10, color = "gray60", margin = margin(0, 0, 3, 0)),
  plot.caption.position = "plot",
  legend.text = element_text(size = 20),
  legend.position = "bottom",
  legend.title = element_text(face = "bold", hjust = 0.5, margin = margin(0,0,3,0), size = 20),
  axis.title.y = element_text(angle = 90, margin = margin(-50,5,20,5), hjust = 0.83, size = 20),
  axis.ticks.y = element_line(color = "gray20", size = 1.5),
  axis.line.x = element_blank(),
  plot.title = element_markdown(size = 30),
  plot.title.position = "plot",
  plot.subtitle = element_text(size = 22),
  axis.text.x = element_text(size = 20, color = 1),
  axis.text.y = element_text(size = 20, color = pal1, angle = 0, margin = margin(0,3,0,0)),
  axis.ticks.x = element_line(color = "grey20", size = 20),
  panel.grid.major.y = element_line(color = c(pal1, "#FFFFFF"), linetype = 1),
  panel.grid.major.x = element_line(color = "grey70", linetype = 3)
)
```
```{r, warning=FALSE}
p <- df %>%
  ggplot(aes(x = yday, y = tmax)) +
  geom_point(aes(color = years), alpha = 0.4, size = 0.4) +
  scale_color_discrete_sequential(palette = "Plasma") +
  scale_x_continuous(breaks = months$yday, labels = months$label, expand = c(.001, .001)) + 
  scale_y_continuous(breaks = c(30, 50, 70, 90, 110), labels = c('30', '50', '70', '90', '110'), expand = c(.001, .001)) + 
  coord_polar() +
  labs(
    y = expression(paste("daily temperature (", degree ~ F, ")"))
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 4, alpha = 1), nrow = 1, title.position = "bottom")
  )
p + 
  labs(
    caption = "**Data:** NOAA | **Visualization:** @jgranna",
    title = "**Daily Maximum Temperatures in Portland, 1939 - 2021**"
  )
  #ggforce::geom_mark_ellipse(  #--- not included; optional ellipse marking ---#
  #  aes(fill = new_record, label = new_record, filter = new_record != "", description = desc), 
  #  alpha = 0.5,color = "00000000", label.colour = "grey20", con.colour = "grey20",
  #  expand = unit(0, "mm"), con.cap = 5, show.legend = FALSE, label.buffer = unit(25, 'mm'),
  #  label.fontsize = c(25, 20), label.family = "osans"
  #)
```

## Smoothed maximum temperatures
The three days in June 2021 are clearly visible at the bottom of the graph, close to July. However, possibly due to the mass of points, a time trend is not easily identifyable. In order to improve this, I plot smoothed functions of the 20-years periods. I choose a GAM smoother, which I chose more or less arbitrarily. However, this is sufficient to be able to identify a time trend. That is, maximum temperatures are generally increasing in Portland, which is of course little surprising.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
pal2 <- sequential_hcl(4, palette = "Red-Blue", rev = T)
p2 <- df %>%
  ggplot(aes(x = yday, y = tmax)) +
  theme(
    axis.text.y = element_text(size = 20, color = pal2, angle = 0, margin = margin(0,3,0,0)),
    panel.grid.major.y = element_line(color = c(pal2, "#FFFFFF"), linetype = 1)
  ) +
  geom_smooth(aes(color = years), alpha = 0.4, size = 0.4, method = "gam", se = F) +
  scale_color_discrete_sequential(palette = "Plasma") +
  scale_x_continuous(breaks = months$yday, labels = months$label, expand = c(.001, .001)) + 
  scale_y_continuous(breaks = c(30, 50, 70, 90), labels = c('30', '50', '70', '90'), limits = c(30, 90), expand = c(.001, .001)) + 
  coord_polar() +
  guides(
    color = guide_legend(override.aes = list(size = 4, alpha = 1), nrow = 1, title.position = "bottom")
  ) 
p2 + labs(
  y = expression(paste("daily temperature (", degree ~ F, ")")),
  caption = "**Data:** NOAA | **Visualization:** @jgranna",
  title = "**Daily Maximum Temperatures in Portland, 1939 - 2021**"
)
```

## Combine both graphs
Finally, for a better overview (and practice), I combine both plots using the `patchwork`- package:
```{r, warning=FALSE, message=FALSE}
library(patchwork)
p + theme(legend.position = "none") + p2 + theme(axis.text.y = element_text(margin = margin(0,6,0,4)), axis.title.y = element_blank()) + scale_y_continuous(position = "right") +
  plot_layout(guides = "collect") +
  plot_annotation(
    caption = "**Data:** NOAA | **Visualization:** @jgranna",
    subtitle = "The left plot shows a scatter plot of maximum temperatures in Portland. The right plot gives the corresponding GAM smooths which makes it easier to visually identify a time trend in the data.", 
    title = "**Daily Maximum Temperatures in Portland, 1939 - 2021**"
  ) &
  theme(plot.caption = element_markdown(margin = margin(0,3,0,0))) 
```

--------------------------------------------------

Full R code available on [Github](https://github.com/jgranna/tidytuesday/tree/main/2021-06-29).

--------------------------------------------------

## Used Packages

- `colorspace`**-package**:  
<https://colorspace.r-forge.r-project.org/articles/colorspace.html>

- `ggtext`**-package**:  
<https://wilkelab.org/ggtext/>

- `patchwork`**-package**:  
<https://patchwork.data-imaginist.com/index.html>

- `showtext`**-package**:  
<https://github.com/yixuan/showtext>

- `tidyverse`**-package**:  
<https://www.tidyverse.org/>
